---
title: "Assignment3 with Python and R"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE}
# Library for transforming objects from Python to R and back
library(reticulate)
# We do not want to get codes printed
knitr::opts_chunk$set(echo = FALSE)
```

```{python sim_setup}

import numpy as np;
import simpy
import random;
import settings
from Hospital import Hospital
# In case you want to save data to to json
import json

def printStatistics(hospital, sample_number, hospital_number):
    patients = hospital.patients

    patients_finished = list(filter(lambda p: p.finished, patients))
    n_finished = len(patients_finished)
    n_patients = len(patients)
    mean_blocking_time = hospital.time_operation_theatre_blocked / n_finished
    mean_queue_at_entrance = hospital.total_queue_at_entrance / n_patients
    utilization_rate_of_operation_theatre = hospital.total_time_operating / settings.SIM_TIME

    print("-" * 20)
    print("sample: ", sample_number)
    print("hospital: ", hospital_number)
    print("Patients came: ", n_patients)
    print("Patients got treated: ", n_finished)
    print("Average time operation theatre was blocked: %.3f" % mean_blocking_time)
    print("Average queue at entrance: %.3f" % mean_queue_at_entrance)
    print("Utilization rate of operation theatre: %.3f" % utilization_rate_of_operation_theatre)

    total_throughput_time = sum([p.end_time - p.start_time for p in patients_finished])
    print("Total throughput time %.3f" % (total_throughput_time))
    print("Average throughput time %.3f" % (total_throughput_time / n_patients))

    print("-" * 20)

def get_data(hospital):
    patients = hospital.patients

    patients_finished = list(filter(lambda p: p.finished, patients))
    n_finished = len(patients_finished)
    n_patients = len(patients)
    mean_blocking_time = hospital.time_operation_theatre_blocked / n_finished
    mean_queue_at_entrance = hospital.total_queue_at_entrance / n_patients
    utilization_rate_of_operation_theatre = hospital.total_time_operating / settings.SIM_TIME

    total_throughput_time = sum([p.end_time - p.start_time for p in patients_finished])

    patients_json = list(map(lambda x: x.to_dict(), patients))

    #"patients": patients_json,
    return {
        "mean_blocking_time": mean_blocking_time,
        "mean_queue_at_entrance": mean_queue_at_entrance,
        "utilization_rate_of_operation_theatre": utilization_rate_of_operation_theatre,
        "total_throughput_time": total_throughput_time
    }


```

## Settings

```{python simulations_settings}
# If you want to use values defined in settings.py file, call settings.N_SAMPLES/settings.SIM_TIME
SIM_TIME = 1000
N_SAMPLES = 20
#PREP_ROOMS = [3,3,4]
#REC_ROOMS = [4,5,5]
N_HOSP = len(settings.CONFIGURATIONS)
CANCELLING_PROB = 0.1
```

## Independent simulations with simPy

```{python independent_simulations}
#initialize results matrix
sim_que = np.zeros( (N_SAMPLES, N_HOSP) ) # average queue time
sim_uti = np.zeros( (N_SAMPLES, N_HOSP) ) # utilization rate
sim_blo = np.zeros( (N_SAMPLES, N_HOSP) )  # average blocking time
#samples are independent, each sample has unique seed (N_SAMPLES*N_HOSP seeds in total)
RANDOM_SEEDS = [*range(N_SAMPLES*N_HOSP)]

for h, config in enumerate(settings.CONFIGURATIONS):
    for sample_i in range(N_SAMPLES):
        random.seed(RANDOM_SEEDS[h*N_SAMPLES + sample_i])
        env = simpy.Environment()
        hospital = Hospital(env, config["n_preparation_rooms"], config["n_recovery_rooms"], CANCELLING_PROB)
        env.run(until = settings.WARM_UP_TIME + SIM_TIME)
        sample_data = get_data(hospital)
        sim_que[sample_i, h] = sample_data["mean_queue_at_entrance"]
        sim_uti[sample_i, h] = sample_data["utilization_rate_of_operation_theatre"]
        sim_blo[sample_i, h] = sample_data["mean_blocking_time"]

# SAVING TO FILE   
#    with open("data.json", mode="w", encoding="utf-8") as f:
#        json.dump(data, f, indent=4)
```

## Calculating results with R
### Differences between independent simulations

```{r, calc_ci}
# Transform Python objects to R. This is possible to replace with json data load
sim_que = py$sim_que
sim_uti = py$sim_uti
sim_blo = py$sim_blo

N_SAMPLES = py$N_SAMPLES
PREP_ROOMS = sapply(py$settings$CONFIGURATIONS, FUN = function(x) x$n_preparation_rooms)
REC_ROOMS = sapply(py$settings$CONFIGURATIONS, FUN = function(x) x$n_recovery_rooms)

# means and standard deviation on the variables
mean_que <- apply(sim_que, 2, mean)
mean_uti <- apply(sim_uti, 2, mean)
mean_blo <- apply(sim_blo, 2, mean)
sd_que <- apply(sim_que, 2, sd)
sd_uti <- apply(sim_uti, 2, sd)
sd_blo <- apply(sim_blo, 2, sd)

# Table of descriptive statistics
descriptives <- data.frame(
    PREP_ROOMS = PREP_ROOMS,
    REC_ROOMS = REC_ROOMS,
    MEAN_QUE = mean_que,
    SD_QUE = sd_que,
    MEAN_UTI = mean_uti,
    SD_UTI = sd_uti,
    MEAN_BLO = mean_blo,
    SD_BLO = sd_blo 
    )
# Table of differences in means between hospitals and 95% confidence interval of
#      the difference based on t-distribution.
# Variable average queue
que_95ci <-data.frame(
    DIFFERENCE = c("1-2", "1-3", "2-3"),
    MEAN_D = c(mean_que[1] - mean_que[2], 
               mean_que[1] - mean_que[3], 
               mean_que[2] - mean_que[3]
               ),
    CI_low = c(t.test(sim_que[,1], sim_que[,2], paired = FALSE)$conf.int[1], 
               t.test(sim_que[,1], sim_que[,3], paired = FALSE)$conf.int[1],
               t.test(sim_que[,2], sim_que[,3], paired = FALSE)$conf.int[1]
    ),
    CI_high = c(t.test(sim_que[,1], sim_que[,2], paired = FALSE)$conf.int[2], 
               t.test(sim_que[,1], sim_que[,3], paired = FALSE)$conf.int[2],
               t.test(sim_que[,2], sim_que[,3], paired = FALSE)$conf.int[2]
    )
)
# Variable utilization rate
uti_95ci <-data.frame(
    DIFFERENCE = c("1-2", "1-3", "2-3"),
    MEAN_D = c(mean_uti[1] - mean_uti[2], 
               mean_uti[1] - mean_uti[3], 
               mean_uti[2] - mean_uti[3]
               ),
    CI_low = c(t.test(sim_uti[,1], sim_uti[,2], paired = FALSE)$conf.int[1], 
               t.test(sim_uti[,1], sim_uti[,3], paired = FALSE)$conf.int[1],
               t.test(sim_uti[,2], sim_uti[,3], paired = FALSE)$conf.int[1]
    ),
    CI_high = c(t.test(sim_uti[,1], sim_uti[,2], paired = FALSE)$conf.int[2], 
               t.test(sim_uti[,1], sim_uti[,3], paired = FALSE)$conf.int[2],
               t.test(sim_uti[,2], sim_uti[,3], paired = FALSE)$conf.int[2]
    )
)

blo_95ci <-data.frame(
    DIFFERENCE = c("1-2", "1-3", "2-3"),
    MEAN_D = c(mean_blo[1] - mean_blo[2], 
               mean_blo[1] - mean_blo[3], 
               mean_blo[2] - mean_blo[3]
               ),
    CI_low = c(t.test(sim_blo[,1], sim_blo[,2], paired = FALSE)$conf.int[1], 
               t.test(sim_blo[,1], sim_blo[,3], paired = FALSE)$conf.int[1],
               t.test(sim_blo[,2], sim_blo[,3], paired = FALSE)$conf.int[1]
    ),
    CI_high = c(t.test(sim_blo[,1], sim_blo[,2], paired = FALSE)$conf.int[2], 
               t.test(sim_blo[,1], sim_blo[,3], paired = FALSE)$conf.int[2],
               t.test(sim_blo[,2], sim_blo[,3], paired = FALSE)$conf.int[2]
    )
)

#Saving for later comparison
descriptives1 <- descriptives
```


```{r print_results, echo = FALSE}
print("Descriptive statistics")
print(round(descriptives,2))
print("95% confidence intervals for differences between hospitals in mean of average entrance:")
print(que_95ci)
print("95% confidence intervals for differences between hospitals in utilization rate:")
print(uti_95ci)
print("95% confidence intervals for differences between hospitals in mean of average entrance:")
print(blo_95ci)
print("Seems that 0 is included in monst of the intervals, two significant differences between hospitals")
```

## Dependent (i.e. contrafactual) hospitals with simPy

```{python dependent_simulations}
# We are using the same settings as earlier
#initialize results matrix
sim_que = np.zeros( (N_SAMPLES, N_HOSP) )
sim_uti = np.zeros( (N_SAMPLES, N_HOSP) )
sim_blo = np.zeros( (N_SAMPLES, N_HOSP) ) 
#samples are dependent, seeds are replicated between hospitals (N_SAMPLE seeds)
RANDOM_SEEDS = [*range(N_SAMPLES)]

for h, config in enumerate(settings.CONFIGURATIONS):
    for sample_i in range(N_SAMPLES):
        random.seed(RANDOM_SEEDS[sample_i])
        env = simpy.Environment()
        hospital = Hospital(env, config["n_preparation_rooms"], config["n_recovery_rooms"], CANCELLING_PROB)
        env.run(until=settings.WARM_UP_TIME + SIM_TIME)
        sample_data = get_data(hospital)
        sim_que[sample_i, h] = sample_data["mean_queue_at_entrance"]
        sim_uti[sample_i, h] = sample_data["utilization_rate_of_operation_theatre"]
        sim_blo[sample_i, h] = sample_data["mean_blocking_time"]

# SAVING TO FILE   
#    with open("data.json", mode="w", encoding="utf-8") as f:
#        json.dump(data, f, indent=4)
```

## Calculating results with R
### Differences between dependent simulations

```{r, calc_ci2}
sim_que = py$sim_que
sim_uti = py$sim_uti
sim_blo = py$sim_blo
PREP_ROOMS = sapply(py$settings$CONFIGURATIONS, FUN = function(x) x$n_preparation_rooms)
REC_ROOMS = sapply(py$settings$CONFIGURATIONS, FUN = function(x) x$n_recovery_rooms)

mean_que <- apply(sim_que, 2, mean)
mean_uti <- apply(sim_uti, 2, mean)
mean_blo <- apply(sim_blo, 2, mean)
sd_que <- apply(sim_que, 2, sd)
sd_uti <- apply(sim_uti, 2, sd)
sd_blo <- apply(sim_blo, 2, sd)

# Calculate differences of original simulated values
d_que = cbind(sim_que[,1] - sim_que[,2], sim_que[,1] - sim_que[,3], sim_que[,2] - sim_que[,3])
d_uti = cbind(sim_uti[,1] - sim_uti[,2], sim_uti[,1] - sim_uti[,3], sim_uti[,2] - sim_uti[,3])
d_blo = cbind(sim_blo[,1] - sim_blo[,2], sim_blo[,1] - sim_blo[,3], sim_blo[,2] - sim_blo[,3])

mean_d_que <- apply(d_que, 2, mean)
mean_d_uti <- apply(d_uti, 2, mean)
mean_d_blo <- apply(d_blo, 2, mean)
#sd_d_que <- apply(d_que, 2, sd)
#sd_d_uti <- apply(d_uti, 2, sd)
#sd_d_blo <- apply(d_blo, 2, sd)

# Table of descriptive statistics
descriptives <- data.frame(
    PREP_ROOMS = PREP_ROOMS,
    REC_ROOMS = REC_ROOMS,
    MEAN_QUE = mean_que,
    SD_QUE = sd_que,
    MEAN_UTI = mean_uti,
    SD_UTI = sd_uti,
    MEAN_BLO = mean_blo,
    SD_BLO = sd_blo 
    )

# Table of descriptive statistics of the differences
descriptives_d <- data.frame(
    PREP_ROOMS = PREP_ROOMS,
    REC_ROOMS = REC_ROOMS,
    MEAN_D_QUE = mean_d_que,
    MEAN_D_UTI = mean_d_uti,
    MEAN_D_BLO = mean_d_blo
    )

# Similar confidence intervals now with paired samples (paired = TRUE)
que_95ci <-data.frame(
    DIFFERENCE = c("1-2", "1-3", "2-3"),
    MEAN_D = descriptives_d$MEAN_D_QUE,
    CI_low = c(t.test(sim_que[,1], sim_que[,2], paired = TRUE)$conf.int[1], 
               t.test(sim_que[,1], sim_que[,3], paired = TRUE)$conf.int[1],
               t.test(sim_que[,2], sim_que[,3], paired = TRUE)$conf.int[1]
    ),
    CI_high = c(t.test(sim_que[,1], sim_que[,2], paired = TRUE)$conf.int[2], 
               t.test(sim_que[,1], sim_que[,3], paired = TRUE)$conf.int[2],
               t.test(sim_que[,2], sim_que[,3], paired = TRUE)$conf.int[2]
    )
)

uti_95ci <-data.frame(
    DIFFERENCE = c("1-2", "1-3", "2-3"),
    MEAN_D = descriptives_d$MEAN_D_UTI,
    CI_low = c(t.test(sim_uti[,1], sim_uti[,2], paired = TRUE)$conf.int[1], 
               t.test(sim_uti[,1], sim_uti[,3], paired = TRUE)$conf.int[1],
               t.test(sim_uti[,2], sim_uti[,3], paired = TRUE)$conf.int[1]
    ),
    CI_high = c(t.test(sim_uti[,1], sim_uti[,2], paired = TRUE)$conf.int[2], 
               t.test(sim_uti[,1], sim_uti[,3], paired = TRUE)$conf.int[2],
               t.test(sim_uti[,2], sim_uti[,3], paired = TRUE)$conf.int[2]
    )
)

blo_95ci <-data.frame(
    DIFFERENCE = c("1-2", "1-3", "2-3"),
    MEAN_D = descriptives_d$MEAN_D_BLO,
    CI_low = c(t.test(sim_blo[,1], sim_blo[,2], paired = TRUE)$conf.int[1], 
               t.test(sim_blo[,1], sim_blo[,3], paired = TRUE)$conf.int[1],
               t.test(sim_blo[,2], sim_blo[,3], paired = TRUE)$conf.int[1]
    ),
    CI_high = c(t.test(sim_blo[,1], sim_blo[,2], paired = TRUE)$conf.int[2], 
               t.test(sim_blo[,1], sim_blo[,3], paired = TRUE)$conf.int[2],
               t.test(sim_blo[,2], sim_blo[,3], paired = TRUE)$conf.int[2]
    )
)

```

```{r print_results2, echo = FALSE}
print("Descriptive statistics for average entrance, utilization rate and blocking time")
print(round(descriptives,2))
print("95% confidence intervals for mean differences between hospitals in average entrance:")
print(que_95ci)
print("95% confidence intervals for mean differences between hospitals in utilization rate:")
print(uti_95ci)
print("95% confidence intervals for differences between hospitals in mean of blocking time:")
print(blo_95ci)
print("Seems that this method is more efficient. There are 4 statistically significant differences between hospitals")
```
 
# Personal twist's effect tot utilization rate
- Our personal twist is 10% probability of operation cancellation during preparation
- Let's repeat independent sample simulation with 0% probability and compare the results

```{python independent_simulations0}
CANCELLING_PROB = 0
#initialize results matrix
sim_uti = np.zeros( (N_SAMPLES, N_HOSP) ) # utilization rate
#samples are independent, each sample has unique seed (N_SAMPLES*N_HOSP seeds in total)
RANDOM_SEEDS = [*range(N_SAMPLES*N_HOSP)]

for h, config in enumerate(settings.CONFIGURATIONS):
    for sample_i in range(N_SAMPLES):
        random.seed(RANDOM_SEEDS[h*N_SAMPLES + sample_i])
        env = simpy.Environment()
        hospital = Hospital(env, config["n_preparation_rooms"], config["n_recovery_rooms"], CANCELLING_PROB)
        env.run(until = settings.WARM_UP_TIME + SIM_TIME)
        sample_data = get_data(hospital)
        sim_uti[sample_i, h] = sample_data["utilization_rate_of_operation_theatre"]
```

## Calculating results with R
### Differences between independent simulations

```{r, uti_compare}
# Transform Python objects to R. This is possible to replace with json data load
sim_uti = py$sim_uti

# means and standard deviation on the variables
mean_uti <- apply(sim_uti, 2, mean)
sd_uti <- apply(sim_uti, 2, sd)

# Table of descriptive statistics
descriptives <- data.frame(
    PREP_ROOMS = PREP_ROOMS,
    REC_ROOMS = REC_ROOMS,
    MEAN_UTI = mean_uti,
    SD_UTI = sd_uti
    )
```

```{r print_results3, echo = FALSE}
print("Descriptive statistics for utilization rate")
print(round(descriptives,2))
print("For comparison original descriptive statistics")
print(round(descriptives1,2))
print("Seems that utilization rate is improving a little bit without the twist.")
```